IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-10-16 paper/formal.tex
Pointer to extended version

We are interested in proving non-interference about many programming
languages.  This requires an appropriate definition of this notion
that is language agnostic, so in this section, we present a few
general definitions for what an information flow control language is
and what non-interference properties it may have.  We also state
theorems about our system which we prove in Section~\ref{sec:proofs}.
In particular, we show that |specLang alpha targetLang|, with an appropriate
scheduler |alpha|, satisfies non-interference~\cite{Goguen82},
without making any reference to properties of |targetLang|.

@@

However, even non-adversarial schedulers are not always safe.
Consider, for example, the sequential scheduling policy |seqf| given in
Figure~\ref{fig:scheduler}.
%
It is easy to show that |specLang seqf targetLang| does not satisfy
TSNI:
%
consider a target language similar to |targetLangML| with an
additional expression terminal |diverge| that denotes a divergent computation,
i.e., |diverge| always reduces to |diverge| and a simple label lattice |{pub,
sec}| such that |canFlowTo2 pub sec|, but |cantFlowTo2 sec pub|.
Consider the following two configurations in this language:
\begin{code}
ic1 = iS ; fullconf 1 sec tS1 (IT(if false then diverge else true)),  fullconf 2 pub tS2 ie
ic2 = iS ; fullconf 1 sec tS1 (IT(if true then diverge else true)),   fullconf 2 pub tS2 ie
\end{code}
These two configurations are |pub|-equivalent, but |ic1| will reduce 
(in two steps) 
to |ic1 = iS ; fullconf 2 pub tS1 (IT true)|, whereas |ic2| will not make
any progress.
%
Suppose that |ie| is a computation that writes to a |pub| channel,\footnote{
Though we do not model labeled channels, extending the calculus with such a
feature is straight forward, see Section~\ref{sec:extensions}.}
then the |sec| task's decision to diverge or not is directly leaked to a
public entity.

>>>>

% We are interested in proving non-interference about many programming
% languages.
%This requires an appropriate definition of this notion
%that is language agnostic, so

In this section, we present a few general and target-language agnostic
definitions for what an information flow control language is and what
non-interference properties it may have.  We also state theorems about our
system. %which we prove in Section~\ref{sec:proofs}.  In particular, we show
that |specLang alpha targetLang|, with an appropriate scheduler |alpha|,
satisfies non-interference~\cite{Goguen82}, without making any reference to
properties of |targetLang|. For technical details, interested readers can refer
to the extended version of this paper~\footnote{\url{http://informationflowcontrol.net/ifc-inside/paper.pdf}}.

@@

However, even non-adversarial schedulers are not always safe.
Consider, for example, the sequential scheduling policy |seqf| given in
Figure~\ref{fig:scheduler}.
%
It is easy to show that |specLang seqf targetLang| does not satisfy
TSNI:
%
consider a target language with a standard branching construct |if e then e else
e|
%similar to |targetLangML|
and an additional expression terminal |diverge| that denotes a divergent
computation, i.e., |diverge| always reduces to |diverge| and a simple label
lattice |{pub, sec}| such that |canFlowTo2 pub sec|, but |cantFlowTo2 sec pub|.
Consider the following two configurations in this language:
\begin{code}
ic1 = iS ; fullconf 1 sec tS1 (IT(if false then diverge else true)),  fullconf 2 pub tS2 ie
ic2 = iS ; fullconf 1 sec tS1 (IT(if true then diverge else true)),   fullconf 2 pub tS2 ie
\end{code}
These two configurations are |pub|-equivalent, but |ic1| will reduce
(in two steps)
to |ic1 = iS ; fullconf 2 pub tS1 (IT true)|, whereas |ic2| will not make
any progress.
%
Suppose that |ie| is a computation that writes to a |pub| channel,\footnote{
Though we do not model labeled channels, extending the calculus with such a
feature is straight forward, see Section~\ref{sec:extensions}.}
then the |sec| task's decision to diverge or not is directly leaked to a
public entity.
